---
title: "Analysis Workflow"
author: "Gemini"
date: "2025-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R Markdown document outlines the complete analysis pipeline. It loads data, performs calculations, runs various analyses, and generates plots.

## Step 1: Load Configuration and Libraries

This chunk loads the main configuration file and all required data loading functions.

```{r load_config}
# Load configuration
source("scripts/config.R")

# Load data loading functions
source("scripts/load_gene_functionality_features.R")
source("scripts/load_dinucleotide_features.R")
source("scripts/load_epigenetic_features.R")
source("scripts/load_gene_functionality_zscores.R")
source("scripts/load_epigenetic_zscores.R")
source("scripts/utils.R")
```

## Step 2: Load Data and Compute Z-Scores

This chunk loads the feature matrices and computes the robust z-scores.

```{r compute_zscores}
# Load and compute z-scores
feature_matrix <- load_gene_functionality_features()
feature_matrix_epigenetic <- load_epigenetic_features()
source("scripts/compute_robust_zscores.R")
```

## Step 3: Run Analysis Scripts

This chunk executes the main analysis scripts, including distance effect, MDS, and PCA analysis.

```{r run_analysis}
# Run analysis scripts
source("scripts/analize_distance_effect.R")
source("scripts/mds_analisys.R")
source("scripts/pca_analysis.R")
```

## Step 4: Generate Plots

This chunk generates all the plots for the analysis, including heatmaps and violin plots.

```{r generate_plots}
# Generate plots
source("scripts/generate_epigenetic_heatmap.R")
source("scripts/generate_intrinsic_heatmap.R")
source("scripts/generate_gene_functionality_heatmap.R")
source("scripts/generate_epigenetic_zscore_jitter_plots.R")
source("scripts/generate_epigenetic_zscore_violin_plots.R")
source("scripts/generate_gene_functionality_zscore_violin_plots.R")
```

## Conclusion

The analysis pipeline has been completed successfully. The results can be found in the `results` directory.

```{r completion_message}
print("Analysis pipeline completed successfully!")
```
